---
title: "YoungminChoi-TTestHW-02-a-Code"
author: "Youngmin Choi"
output: 
  html_document:
      df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective/Goal
- To compare 20 colon cancer patients that relapsed with 20 patients that did not. 
- To analyze which genes are differentially changed between two groups. 
- This will help to understand affecting factors for such relapse. 

## Analysis Steps 

### Step 0 - Check Clinical and Molecular Data 
- Clinical data ("CRC_PILOT_clinical_data_HIDS.tsv")
    - 40 rows confirmed. (20 relapse, 20 non-relapse confirmed.)
    - Rows are patient (subject) ID & columns are attributes. 
- Molecular data (Gene data, "CRC_PILOT_withGeneAnno.tsv")
    - Total 80 rows (40 tumor samples, 40 normal samples)
    - Rows are sample ID & columns are gene annotations.  


### Step 1 - Read in Data
```{r}
clinical_data <- read.table(file = "input/CRC_PILOT_clinical_data_HIDS.tsv",
                      sep="\t",
                      row.names = 1,
                      header = T, 
                      stringsAsFactors = F)
molecular_data <- read.table(file = "input/CRC_PILOT_withGeneAnno.tsv",
                      sep="\t",
                      row.names = 1,
                      header = T, 
                      stringsAsFactors = F)
head(clinical_data[,c(29:32)])
head(molecular_data)

molecular_data_t <- as.data.frame(t(molecular_data)) # Transposed the data so that the gene annotations are row and sample IDs are columns. 

knitr::kable(table(clinical_data$RECURRENCE_ANY))
```

### Step 2 - Clean/Filter Data
```{r}
dim(clinical_data)

library(dplyr)
tumor_samples<-molecular_data_t %>% select(contains("Tumor")) # Filtered only tumor samples from molecular data (transposed). 40 samples confirmed! 

# Change tumor_samples column name so that it is identical to clinical_data ID. 
column_name<- colnames(tumor_samples)
column_name
split_names <- strsplit(column_name, "_")
new_name <- sapply(split_names, function(x) x[2])
colnames(tumor_samples) <- new_name
```


### Step 3 - Identify the Groups to be Compared
```{r}
# Baseline group (no-relapse group) row numbers
baseline_group <- which(clinical_data$RECURRENCE_ANY == "NO") # 20 samples
head(baseline_group)
# Baseline group data from clinical_data. 
clin_baseline <- clinical_data[baseline_group,]
baseline_id <- rownames(clin_baseline) # baseline sample id 

# Comparison group (relapse group) row numbers
comparison_group <- which(clinical_data$RECURRENCE_ANY == "YES") # 20 samples
head(comparison_group)
# Baseline group data from clinical_data. 
clin_comparison <- clinical_data[comparison_group,]
comparison_id <- rownames(clin_comparison)

# Molecular data subset 
# Baseline group molecular data
molecular_baseline <- tumor_samples[,baseline_id] # 20 variables, 17881 rows
# Comparison group molecular data
molecular_comparison <- tumor_samples[,comparison_id] # 20 variables, 17881 rows


```

### Step 4 - Sanity Check
```{r}
# See if sample IDs match in clinical data and molecular data
rownames(clin_baseline) == colnames(molecular_baseline)
rownames(clin_comparison) == colnames(molecular_comparison)

# Check by .tsv file comparison. 
write.table(rownames(clin_baseline), file="output/YoungminChoi_ClinBaseIDs.tsv", sep="\t", quote=F)
write.table(rownames(clin_comparison), file="output/YoungminChoi_ClinCompIDs.tsv", sep="\t", quote=F)

write.table(colnames(molecular_baseline), file="output/YoungminChoi_GeneExpBaseIDs.tsv", sep="\t", quote=F)
write.table(colnames(molecular_comparison), file="output/YoungminChoi_GeneExpCompIDs.tsv", sep="\t", quote=F)

write.table(rownames(molecular_baseline), file="output/YoungminChoi_FeatureIDs.tsv")
```

### Step 5 - Prep Data for T-test
```{r}
knitr::kable(head(molecular_baseline[1:5,1:4]))
knitr::kable(head(molecular_comparison[1:5,1:4]))

```

### Step 6 - Call Function for T-test 
```{r}
source("fnTTest.R")

results1 = fnTTest(baseGroup = molecular_baseline, 
                   compGroup = molecular_comparison, 
                   testName = "TTest_HW04", 
                   baseGroupName = "No_relapse",
                   compGroupName =  "Relapsed",
                   folderName = "output")

```

### Step 6-II - Further Result Analysis

```{r}
ttestResults <- read.csv(file = "output/TTest_HW04_Relapsed_(Comp).vs._No_relapse_(Base).TTest.csv")

ttestResultsSorted <- dplyr::arrange(ttestResults, Pvalue)


whichSig <- which(ttestResultsSorted$Pvalue <= 0.05) 

ttestResultsSig <- ttestResultsSorted[whichSig, ] #719 rows 

### Export short listed results
write.table(x = ttestResultsSig, 
            file = "output/YoungminChoi-TTestHW-04-Output.csv", 
            quote = F, sep = "\t", row.names=FALSE)

gene_names <- ttestResultsSig[,'Feature']
gene_top20<- gene_names[1:20]
gene_top20


### Export list of gene names
write.table(x = gene_top20, 
            file = "output/YoungminChoi-TTestHW-05-Top20Genes.tsv", 
            quote = F, sep = "\t")



```
